---
title: "`tsgx` Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{`tsgx` Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## About the package

**`tsgx`** stands for "**table summary generator**." This package is designed to facilitate generation of statistical summary tables with ease. It also adheres to the `tidyverse` specifications.

The package allows you to:

- generate frequency tables, cross-tabulations (2-way table or more);
- extract multiple-letter response variable from survey data;
- include percent distribution (which is default) in the generated tables;
- specify whether the 'percent to total' is computed by row (default) or by column;
- export Excel file with default formatting/styling which is can also be customized.

## Installation

You may install the `tsgx` package either from GitHub or R-CRAN.

```{r setup, eval=F}
# Install devtools if not yet installed in your machine
if(!('devtools' %in% installed.packages()[,'Package'])){
   install.packages('devtools')
}

# Install the package from GitHub
devtools::install_github('yng-me/tsgx')

# Install via R-CRAN
install.packages('tsgx')

```

Then load the package after installation.

```{r}
library(tsgx)
```


## `tsgx` core functions

### 1. **`generate_frequency`**

This function allows you to generates a frequency distribution table (marginal table) of a categorical variable `x` specified in its second argument. It returns five (5) columns by default if `x_group` is not specified. These include (1) categories of `x`, (2) frequency of each category, (3) percent to total, (4) cumulative frequency, and (5) cumulative percent to total.

```{r, eval=F}
generate_frequency(
  .data,
  x,
  label_stub = get_config('label_stub'),
  sort_frequency = FALSE,
  include_total = TRUE,
  include_cumulative = TRUE,
  exclude_zero_value = FALSE
)
```


**Parameters**: 

<table>
  <tr>
    <td>`.data`</td>
    <td>
      <b>Required</b>. 
      A data frame, data frame extension (e.g. a tibble), a lazy data frame (e.g. from dbplyr or dtplyr), or Arrow data format.
    </td>
  </tr>
  <tr>
    <td>`x`</td>
    <td><b>Required</b>. Variable to be used as categories.</td>
  </tr>
  <tr>
    <td>`label_stub`</td>
    <td>Stubhead label or label for `x`.</td>
  </tr>
  <tr>
    <td>`sort_frequency`</td>
    <td>Whether to sort the output. If set to `TRUE`, the frequency will be sorted in descending order.</td>
  </tr>
  <tr>
    <td>`include_total`</td>
    <td>Whether to include row total.</td>
  </tr>
   <tr>
    <td>`include_cumulative`</td>
    <td>Whether to cumulative frequencies.</td>
  </tr>
   <tr>
    <td>`exclude_zero_value`</td>
    <td>Whether to drop categories with zero (0) values</td>
  </tr>
</table>


**Example 1.1**: Basic usage

```{r, warning=F}
library(palmerpenguins)
library(dplyr)
library(gt)
```

**Example 1.2**: Add grouping variable and define label for `x`

```{r, warning=F}
penguins |> 
  generate_frequency(
    x = sex, 
    label_stub = 'Sex'
  )
```

**Example 1.3**: Add grouping variable, and exclude column total

```{r, warning=F}
penguins |> 
  group_by(species) |>
  generate_frequency(
    x = sex, 
    include_total = FALSE
  )
```

**Example 1.4**: Exclude cumulative values and sort the output by frequency

```{r, warning=F}
penguins |> 
  generate_frequency(
    x = species, 
    label_stub = 'Species', 
    sort_frequency = TRUE, 
    include_cumulative = FALSE
  )
```

**Example 1.5**: Exclude cumulative values and define multiple grouping variables

```{r, warning=F}
dplyr::starwars |> 
  group_by(skin_color, gender) |> 
  generate_frequency(
    x = sex, 
    label_stub = 'Sex', 
    include_cumulative = FALSE
  )
```


### 2. **`generate_crosstab`** 

**`generate_crosstab`** extends the functionality of `generate_frequency` by allowing you to generate cross-tabulations of two (2) or more categorical variables.

```{r, eval=F}
generate_crosstab(
  .data,
  x,
  ...,
  total_by_col = FALSE,
  total_option = 'default',
  include_frequency = TRUE,
  include_proportion = TRUE,
  include_column_total = TRUE,
  include_row_total = TRUE,
  include_subtotal = FALSE,
  include_zero_value = FALSE,
  convert_to_percent = TRUE,
  decimal_precision = NULL,
  label_stub = get_config('label_stub'),
  label_total = 'Total',
  label_subtotal = NULL,
  weights = NULL,
  names_separator = '>'
) 
```

**Parameters**: 

<table>
  <tr>
    <td>`.data`</td>
    <td>
      <b>Required</b>. 
      A data frame, data frame extension (e.g. a tibble), a lazy data frame (e.g. from dbplyr or dtplyr), or Arrow data format.</td>
  </tr>
  <tr>
    <td>`x`</td>
    <td><b>Required</b>. Variable to be used as categories.</td>
  </tr>
  <tr>
    <td>`...`</td>
    <td>Tidy-select column names.</td>
  </tr>
  <tr>
    <td>`total_by_col`</td>
    <td>Accepts `TRUE` | `FALSE`. Whether to apply the sum columnwise if `TRUE` or rowwise `FALSE`. Default is `FALSE`.</td>
  </tr>
  <tr>
    <td>`total_option`</td>
    <td>*[CURRENTLY IGNORE]*. For future implementation.</td>
  </tr>
  <tr>
    <td>`include_frequency`</td>
    <td>Whether to include frequency columns. Default is `TRUE`</td>
  </tr>
  <tr>
    <td>`include_proportion`</td>
    <td>Whether to include proportion/percentage columns. Default is `TRUE`</td>
  </tr>
  <tr>
    <td>`include_column_total`</td>
    <td>Whether to include column total. Default is `TRUE`.</td>
  </tr>
  <tr>
    <td>`include_row_total`</td>
    <td>Whether to include row total. Default is `TRUE`</td>
  </tr>
  <tr>
    <td>`include_subtotal`</td>
    <td>Whether to include subtotal. Default is `FALSE`</td>
  </tr>
  <tr>
    <td>`include_zero_value`</td>
    <td>Whether to drop columns with zero (0) values. Default is `FALSE`</td>
  </tr>
  <tr>
    <td>`convert_to_percent`</td>
    <td>Whether to format to percent or proportion.</td>
  </tr>
  <tr>
    <td>`format_precision`</td>
    <td>*[Not yet implemented]* Specify the precision of rounding the percent or proportion. Default is `2`.</td>
  </tr>
  <tr>
    <td>`label_stub`</td>
    <td>Stubhead label or label for `x`.</td>
  </tr>
  <tr>
    <td>`label_total`</td>
    <td>Whether to rename the column total.</td>
  </tr>
  <tr>
    <td>`weight`</td>
    <td>*[CURRENTLY IGNORE]*. For future implementation.</td>
  </tr>
  <tr>
    <td>`names_separator`</td>
    <td>A character string that defines the column separator to be used to show table hierarchy.</td>
  </tr>
</table>

**Example 2.1**: Basic usage

```{r}
penguins |> 
  generate_crosstab(
    species, 
    sex
  )
```

**Example 2.2**: Percent/proportion total by column

```{r}
penguins |> 
  generate_crosstab(
    species, 
    sex,
    total_by_col = TRUE
  )
```

**Example 2.3**: Exclude frequencies

```{r}
penguins |> 
  generate_crosstab(
    species, 
    sex,
    include_frequency = F
  )
```

**Example 2.4**: Exclude percentages/proportions

```{r}
penguins |> 
  generate_crosstab(
    species, 
    sex,
    include_proportion = F
  )
```

**Example 2.5**: Add row grouping variable and exclude percentages/proportions

```{r}
penguins |> 
  group_by(island) |>
  generate_crosstab(
    species, 
    sex, 
    include_proportion = F
  )
```

**Example 2.6**: Add column grouping variable, exclude frequencies, and convert_to_percent set to `FALSE`.

```{r}
penguins |> 
  group_by(island) |>
  generate_crosstab(
    species, 
    sex, 
    convert_to_percent = F,
    include_frequency = F
  )
```


### 3. **`generate_multiple_response`**

This function allows you to generate summary table from a multiple response category.

```{r, eval=F}
generate_multiple_response(
  .data,
  x,
  ...,
  value_to_count = 1,
  include_frequency = TRUE,
  include_proportion = TRUE,
  convert_to_percent = TRUE,
  format_precision = 2,
  label_stub = get_config('label_stub'),
  label_total = NULL,
  names_separator = '>',
  recode = TRUE,
  recode_variable_position = 5,
  clean_name = TRUE
) 
```


**Parameters**: 

<table>
  <tr>
    <td>`.data`</td>
    <td>
      <b>Required</b>. 
      A data frame, data frame extension (e.g. a tibble), a lazy data frame (e.g. from dbplyr or dtplyr), or Arrow data format.</td>
  </tr>
  <tr>
    <td>`x`</td>
    <td><b>Required</b>. Variable to be used as categories.</td>
  </tr>
  <tr>
    <td>`...`</td>
    <td>tidyselect columns. If only one variable is specified, it will be treated as letter-coded response; otherwise, it will be parsed as columns with binary-coded response.</td>
  </tr>
  <tr>
    <td>`value_to_count`</td>
    <td>Value used as basis for counting the frequencies.</td>
  </tr>
  <tr>
    <td>`include_frequency`</td>
    <td>Whether to include frequency columns.</td>
  </tr>
  <tr>
    <td>`include_proportion`</td>
    <td>Whether to include proportion/percentage columns.</td>
  </tr>
  <tr>
    <td>`convert_to_percent`</td>
    <td>Whether to format to percent or proportion.</td>
  </tr>
  <tr>
    <td>`format_precision`</td>
    <td>*[Not yet implemented]* Specify the precision of rounding the percent or proportion. Default is `2`.</td>
  </tr>
  <tr>
    <td>`label_stub`</td>
    <td>Stubhead label or label for `x`.</td>
  </tr>
  <tr>
    <td>`label_total`</td>
    <td>Whether to rename the column total.</td>
  </tr>
  <tr>
    <td>`recode`</td>
    <td>Whether to recode the variable name first.</td>
  </tr>
  <tr>
    <td>`recode_variable_position`</td>
    <td>Indicate the position of letter of variable that uniquely identifies each variable. Default value is `5`</td>
  </tr>
  <tr>
    <td>`clean_name`</td>
    <td>Whether to output a clean column name.</td>
  </tr>
  <tr>
    <td>`names_separator`</td>
    <td>A character string that defines the column separator to be used to show table hierarchy.</td>
  </tr>
</table>


**Example 3.1**: Basic usage (extract multiple-letter response)

```{r, eval=F}

df <- data.frame(
  group = c("A1", "A2", "A1", "A1", "A2", "A2"),
  category = c("G1", "G1", "G2", "G1", "G2", "G1"),
  response = c("AB", "AC", "B", "ABC", "AB", "C"),
  A = c(1, 1, 0, 1, 1, 0),
  B = c(1, 0, 1, 1, 1, 0),
  C = c(0, 1, 0, 1, 0, 1)
) 

df |> 
  group_by(group) |>
  generate_multiple_response(category, response)
```

**Example 3.1**: Basic usage (wide format multiple response)

```{r, eval=F}
df |> generate_multiple_response(category, matches('^[A-C]$'), recode_variable_position = 1)
```
